---
title: "The correlation of diabetes and obesity with physical inactivity in the US in 2017"
author: "Sam (Cheng-Hsiang) Lu"
date: "`r Sys.Date()`"
output: 
        pdf_document: default
        html_document:
          toc: TRUE
          toc_float: TRUE
---

<br>

# **Introduction**

In this project, I would like to see the correlation of diabetes and obesity with physical inactivity in the US in 2017. Thus, I download two datasets which are talking about "diagnosed diabetes among adults aged >=18 years" and "Obesity among adults aged >=18 years" in the US in 2017 from the CDC. They include estimates for the 500 largest US cities and approximately 28,000 census tracts within these cities. I also download one dataset which is the "Prevalence of Self-Reported Physical Inactivity Among US Adults by State and Territory, 2017–2020".

<br>

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
library(data.table)
library(tidyverse)
library(dplyr)
library(plotly)
library(DT)
library(knitr)
library(lubridate)
library(leaflet)
library(dtplyr)
library(ggplot2)
library(stringr)
library("RSocrata")

# Initialize code chunk options
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = TRUE,
  cache = FALSE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px")
```

# **Methods**

## Read in the data by API

I used API method to obtain my datasets from CDC. First, you have to create an account with password. Then, you have to apply for a free app token. Last, copy your API Endpoint. Both datasets contain 27 columns and 29,006 rows. 

Here are my datasets links:

`https://chronicdata.cdc.gov/500-Cities-Places/500-Cities-Obesity-among-adults-aged-18-years/bjvu-3y7d`

`https://chronicdata.cdc.gov/500-Cities-Places/500-Cities-Diagnosed-diabetes-among-adults-aged-18/cn78-b9bj`

```{r API_dia, message=FALSE, echo=FALSE, warning=FALSE}
dia <- read.socrata(
  "https://chronicdata.cdc.gov/resource/cn78-b9bj.json?year=2017",
  app_token = "bEkVW73ASzmTkZ9riAtf2YS5c",
  email     = "clu74108@usc.edu",
  password  = "Samuelsunny0325!"
)
dia <- as.data.table(dia)
write.csv(dia,"./data/dia.csv", row.names = F)
```

```{r API_obe, message=FALSE, echo=FALSE, warning=FALSE}
obe <- read.socrata(
  "https://chronicdata.cdc.gov/resource/bjvu-3y7d.json?year=2017",
  app_token = "bEkVW73ASzmTkZ9riAtf2YS5c",
  email     = "clu74108@usc.edu",
  password  = "Samuelsunny0325!"
)
obe <- as.data.table(obe)
write.csv(obe,"./data/obe.csv", row.names = F)
```

<br>

The Physical Inactivity dataset contains 3 columns and 53 rows. 

This is the URL to download it:

`https://www.cdc.gov/physicalactivity/data/inactivity-prevalence-maps/tables/2020/1-self-reported.csv`
```{r message=FALSE, echo=FALSE, warning=FALSE}
if (!file.exists("1-self-reported.csv"))
  download.file(
    url = "https://www.cdc.gov/physicalactivity/data/inactivity-prevalence-maps/tables/2020/1-self-reported.csv",
    destfile = "1-self-reported.csv",
    method   = "libcurl",
    timeout  = 60
  )
physical_inactivity <- fread("1-self-reported.csv")
write.csv(physical_inactivity,"./data/1-self-reported.csv", row.names = F)
```

<br>

## Select columns

From CDC datasets, I select **data_value(%), populationCount, stateabbr, statedesc(state name), city_name, geolocation.latitude, and geolocation.longitude** total 7 columns. From the Physical Inactivity dataset, I select **state and prevalence** 2 columns.

```{r message=FALSE, echo=FALSE, warning=FALSE}
dia_mini <- dia[, c(2, 3, 14, 17, 19, 20, 23)]
obe_mini <- obe[, c(2, 3, 14, 17, 19, 20, 23)]
ph_inact <- physical_inactivity[, c(1, 2)]
```

<br>

## Change column names
I change my column names in order to easily understand. In diabetes and obesity datasets, I have  **diabetes_percentage/obesity_percentage, populationCount, state_abbr, state_name, city_name, lat, and lon**. In the physical inactivity dataset, I have **state_name and physical_inactivity_prevalence**.

```{r message=FALSE, echo=FALSE, warning=FALSE}
colnames(dia_mini)[1] <- "diabetes_percentage"
colnames(dia_mini)[2] <- "populationCount"
colnames(dia_mini)[3] <- "state_abbr"
colnames(dia_mini)[4] <- "state_name"
colnames(dia_mini)[5] <- "city_name"
colnames(dia_mini)[6] <- "lat"
colnames(dia_mini)[7] <- "lon"

colnames(obe_mini)[1] <- "obesity_percentage"
colnames(obe_mini)[2] <- "populationCount"
colnames(obe_mini)[3] <- "state_abbr"
colnames(obe_mini)[4] <- "state_name"
colnames(obe_mini)[5] <- "city_name"
colnames(obe_mini)[6] <- "lat"
colnames(obe_mini)[7] <- "lon"

colnames(ph_inact)[1] <- "state_name"
colnames(ph_inact)[2] <- "physical_inactivity_prevalence"
```

```{r message=FALSE, echo=FALSE, warning=FALSE}
ph_inact$state_name <- gsub("DC", "District of C", ph_inact$state_name)
ph_inact$state_name <- gsub("NewHampshire", "New Hampshire", ph_inact$state_name)
ph_inact$state_name <- gsub("NewJersey", "New Jersey", ph_inact$state_name)
ph_inact$state_name <- gsub("NewMexico", "New Mexico", ph_inact$state_name)
ph_inact$state_name <- gsub("NewYork", "New York", ph_inact$state_name)
ph_inact$state_name <- gsub("NorthCarolina", "North Carolin", ph_inact$state_name)
ph_inact$state_name <- gsub("NorthDakota", "North Dakota", ph_inact$state_name)
ph_inact$state_name <- gsub("RhodeIsland", "Rhode Island", ph_inact$state_name)
ph_inact$state_name <- gsub("SouthCarolina", "South Carolin", ph_inact$state_name)
ph_inact$state_name <- gsub("SouthDakota", "South Dakota", ph_inact$state_name)
ph_inact$state_name <- gsub("WestVirginia", "West Virginia", ph_inact$state_name)
```

<br>

## Remove non-numeric value

This step, I remove the physical inactivity prevalence non-numeric value in the Physical Inactivity dataset.

```{r message=FALSE, echo=FALSE, warning=FALSE}
ph_inact <- ph_inact[!is.na(as.numeric(ph_inact$physical_inactivity_prevalence)), ]
```

<br>

## Merge two datasets

Merge CDC two datasets by **state_abbr, populationCount, state_name, city_name, lat, and lon**. Then I merge the Physical Inactivity dataset with them.

```{r merge-datatables, message=FALSE, echo=FALSE, warning=FALSE}
merged_1 <- 
  merge(
  # Data
  x     = dia_mini,      
  y     = obe_mini, 
  # List of variables to match
 by = c("state_abbr","populationCount", "state_name", "city_name", "lat", "lon"),
  # keep everything!
  all.x = TRUE     
  ) 
```

```{r message=FALSE, echo=FALSE, warning=FALSE}
merged <- 
  merge(
    # Data
    x     = merged_1,      
    y     = ph_inact, 
    # List of variables to match
    by = c("state_name"),
    # keep everything!
    all.x = TRUE     
  ) 
```
 
 <br>
 
## Remove duplicates

```{r message=FALSE, echo=FALSE, warning=FALSE}
dim(merged)
```

My row number increased to 30,008 so I have to remove duplicates.

```{r message=FALSE, echo=FALSE, warning=FALSE}
merged[, n := 1:.N, by = .(state_abbr, state_name, city_name, lat, lon, physical_inactivity_prevalence)]
merged <- merged[n == 1,][, n := NULL]

dim(merged)
```

After removing duplicates, my rows shrink from 30,008 to 28,505.

<br>

## Convert chr into num

In this step, I just convert character variables into numeric variables.

```{r message=FALSE, echo=FALSE, warning=FALSE}
#str(merged)
```
```{r message=FALSE, echo=FALSE, warning=FALSE}
merged$lat <- as.numeric(merged$lat)
merged$lon <- as.numeric(merged$lon)
merged$diabetes_percentage <- as.numeric(merged$diabetes_percentage)
merged$populationCount <- as.numeric(merged$populationCount)
merged$obesity_percentage <- as.numeric(merged$obesity_percentage)
merged$physical_inactivity_prevalence <- as.numeric(merged$physical_inactivity_prevalence)
```

<br>

## Check NAs

In my merged dataset, there are only 2.7% NAs values in columns of **diabetes_percentage and obesity_percentage**. Therefore, I'm going to remove them.

```{r message=FALSE, echo=FALSE, warning=FALSE}
#summary(merged$diabetes_percentage)
#summary(merged$obesity_percentage)
```

```{r message=FALSE, echo=FALSE, warning=FALSE}
mean(is.na(merged$diabetes_percentage))
mean(is.na(merged$obesity_percentage))
```

```{r message=FALSE, echo=FALSE, warning=FALSE}
merged <-merged[!is.na(merged$diabetes_percentage),]
merged <-merged[!is.na(merged$obesity_percentage),]
```

```{r message=FALSE, echo=FALSE, warning=FALSE}
#summary(merged$diabetes_percentage)
#summary(merged$obesity_percentage)
```

<br>

## Add regions

Last, I create a new column contain Northeast, Northwest, Southwest, and Southeast four different regions.

```{r message=FALSE, echo=FALSE, warning=FALSE}
# Add regions
merged[, region := fifelse(lon >= -98 & lat > 39.71, "NE",
                fifelse(lon < -98 & lat > 39.71, "NW",
                fifelse(lon < -98 & lat <= 39.71, "SW","SE")))
   ]
#table(merged$region)
```

<br>

# **Results**

## Leaflet {.tabset}

```{r pal_dia, message=FALSE, echo=FALSE, warning=FALSE}
pal_dia <- colorNumeric(c('darkblue','goldenrod','darkred'), domain=merged$diabetes_percentage)
```

```{r p1_leaflet, message=FALSE, echo=FALSE, warning=FALSE}
# Diabetes percentage in the US
p1_leaflet <- leaflet() %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(data = merged,
             lat=~lat,lng=~lon,
             label = ~paste0(round(diabetes_percentage,2)), color = ~ pal_dia(diabetes_percentage),
             opacity = 0.5, fillOpacity = 1, radius = 50) %>%
  # Legend
  addLegend('bottomleft', pal=pal_dia, values=merged$diabetes_percentage,
             title='Diabetes percentage', opacity=1)
```

```{r pal_obe, message=FALSE, echo=FALSE, warning=FALSE}
pal_obe <- colorNumeric(c('darkblue','goldenrod','darkred'), domain=merged$obesity_percentage)
```

```{r p2_leaflet, message=FALSE, echo=FALSE, warning=FALSE}
# Obesity percentage in the US
p2_leaflet <- leaflet() %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(data = merged,
             lat=~lat,lng=~lon,
             label = ~paste0(round(obesity_percentage,2)), color = ~ pal_obe(obesity_percentage),
             opacity = 0.5, fillOpacity = 1, radius = 50) %>%
  # And a pretty legend
  addLegend('bottomleft', pal=pal_obe, values=merged$obesity_percentage,
             title='Obesity percentage', opacity=1)
```

### Diabetes percentage in the US

```{r echo=FALSE, message=FALSE, echo=FALSE, warning=FALSE}
p1_leaflet
```

### Obesity percentage in the US

```{r echo=FALSE, message=FALSE, echo=FALSE, warning=FALSE}
p2_leaflet
```

## {-}
From the Leaflet, the legend shows the degree of the diabetes percentage. The red color means higher percentage of diabetes. I see there are more orange dots in the NE region and SE region from the plot of diabetes percentage. From the second plot of obesity percentage, we can see that there are more orange dots on the east-side than on the west-side.

<br>

## Boxplots {.tabset}

```{r message=FALSE, echo=FALSE, warning=FALSE}
merged$region <- factor(merged$region, levels=c("NE", "SE", "NW", "SW"))
```

```{r p1_box, message=FALSE, echo=FALSE, warning=FALSE}
p1_box <- merged[!is.na(diabetes_percentage)][!is.na(region)] %>% 
  plot_ly(x = ~region, y= ~diabetes_percentage,
        type = 'box', mode = 'markers', color = ~region,
        hoverinfo = 'text',
        text = ~paste( paste(" State name: ", state_name, sep=""),
                       paste(" Region: ", region, sep=""),
                       paste("City name: ", city_name, sep=""),
                       paste(" Diabetes percentage: ", diabetes_percentage, sep=""), 
                       sep = "<br>")) %>%
  layout(title = "Diabetes percentage in different regions",
         xaxis = list(title = "Regions"), 
         yaxis = list(title = "Diabetes percentage"),
         hovermode = "compare")
```

```{r p2_box, message=FALSE, echo=FALSE, warning=FALSE}
p2_box <- merged[!is.na(obesity_percentage)][!is.na(region)] %>% 
  plot_ly(x = ~region, y= ~obesity_percentage,
        type = 'box', mode = 'markers', color = ~region,
        hoverinfo = 'text',
        text = ~paste( paste(" State name: ", state_name, sep=""),
                       paste(" Region: ", region, sep=""),
                       paste("City name: ", city_name, sep=""),
                       paste(" Obesity percentage: ", obesity_percentage, sep=""), 
                       sep = "<br>")) %>%
  layout(title = "Obesity percentage in different regions",
         xaxis = list(title = "Regions"), 
         yaxis = list(title = "Obesity percentage"),
         hovermode = "compare")
```

### Diabetes percentage

```{r message=FALSE, echo=FALSE, warning=FALSE}
p1_box
```

### Obesity percentage

```{r message=FALSE, echo=FALSE, warning=FALSE}
p2_box
```

## {-}

Now, let see the boxplot, the x-axis shows 4 regions: Northeast, Southeast, Northwest, and Southwest. On the y-axis shows the percentage of diabetes or obesity.

From the boxplot of diabetes percentage, there is a max diabetes percentage in the NE region, and the NE region and the SE region have a similar median diabetes percentage. The NW region has the lowest median diabetes percentage. In this plot, the east-side regions' median diabetes percentage is higher than the west-side regions'.

From the boxplot of obesity percentage, the east-side regions' median obesity percentage is also higher than the west-side regions'. Among all regions, the SE region has the highest median obesity percentage and the SW region has the lowest median obesity percentage.

<br>



## Scatter plots {.tabset}

```{r message=FALSE, echo=FALSE, warning=FALSE}
merged_median <- merged[, .(
    obe_median      = median(obesity_percentage, na.rm=TRUE),
    dia_median      = median(diabetes_percentage, na.rm=TRUE),
    physical_inactivity_prevalence = physical_inactivity_prevalence
    ),
    by = c("state_abbr", "state_name", "region")
    ]
```

```{r message=FALSE, echo=FALSE, warning=FALSE}
merged_median_uni <- unique(merged_median)
```

```{r message=FALSE, echo=FALSE, warning=FALSE}
scatter_p <- merged_median_uni %>% 
  plot_ly(x = ~obe_median, y = ~dia_median,
          type = 'scatter', mode = 'markers', color = ~state_abbr,
          size = ~physical_inactivity_prevalence, sizes = c(5, 70), marker = list(sizemode='diameter', opacity=0.5),
          hoverinfo = 'text',
          text = ~paste( paste(state_name, ":", sep=""), 
                  paste(" State_abbr: ", state_abbr, sep=""), 
                  paste(" Region: ", region, sep=""), 
                  paste(" Obesity percentage: ", obe_median, sep=""), 
                  paste(" Diabetes percentage: ", dia_median, sep=""),
                  paste(" Physical Inactivity Prevalence: ", physical_inactivity_prevalence, sep=""),
                  sep = "<br>")) %>%
  layout(title = "Obesity percentage vs. Diabetes percentage with Physical Inactivity Prevalence",
         xaxis = list(title = "Obesity percentage"), 
         yaxis = list(title = "Diabetes percentage"),
         hovermode = "compare")

scatter_p
```

In this scatter plot, I select each state's median of obesity percentage and diabetes percentage with their physical inactivity prevalence. We can see that there is a positive correlation between obesity and diabetes rates. We can also see the physical inactivity prevalence by the size of dots. I cannot tell the direct correlation with the obesity or diabetes rates. However, it looks like there are bigger dots on the top right corner and smaller dots on the bottom left corner.

<br>


# **Conclusion**

## Question 1: How are the distribution of diabetes and obesity percentages in the US?

From the leaflet,  first we can see there are more orange dots on the NE and SE regions. From the box plot, the median of diabetes percentage looks equally high in the NE and SE regions. The median of obesity percentage looks higher in the SE region than in the NE region. Besides, we can also see there are higher diabetes and obesity percentages on the east-side than on the west-side.

## Question 2: Is there any correlation between diabetes and obesity? If so, is there any correlation with physical inactivity prevalence?

From the scatter plot, we can see that there is a positive correlation between obesity and diabetes rates by states. As for the correlation with physical inactivity prevalence, I don't see a strong correlation with physical inactivity prevalence. However, I can tell that smaller dots are on the bottom left corner and bigger dots are on the top right corner.

<br>

# **Discussion**

I would like to add more variables to my dataset in order to make more interesting plots. For example, I want to add different years and time to check if there are any changes of diabetes and obesity percentages in each state. Adding each state's median income level , average education level, and so on, to see whether diabetes and obesity percentages would be affected by them.

